---
title: "ARA_vs_15N2_data_analysis_2025"
author: "Paula Gomez - SCU"
output:
  pdf_document: default
  html_document: default
  word_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Acetylene Reduction Assay

## Read the data sets from ARA:
```{r, results='hide',echo=FALSE, warning=FALSE, message=FALSE}
library(dplyr)
library(ggplot2)
library(purrr)
library(emmeans)
library(car)
library(MASS)
library(tidyr)
library(permute)
library(lattice)
library(rstatix)
library(patchwork)
library(DHARMa)
```

## Read the data sets from ARA:
```{r}
ARA <- read.csv("Data/ARA_data_publication.csv", stringsAsFactors = TRUE)

```

- Research question How does time and treatment affect N fixation using ARA
- Variables - Treatment, Time, N fixation (ngN/g/day)

## View the data and structure the data

```{r}
ARA$incubation_time_days <- as.factor(ARA$incubation_time_days)
ARA_reduced <- ARA %>% dplyr::select(ID, Treatment, Concentration_ppm, incubation_time_days, ngN_g_day)
ARA_reduced <- ARA_reduced %>% 
  rename(
    Nfix_rate_ngN_g_day = ngN_g_day,
  )

ARA_reduced$Treatment <- factor(ARA_reduced$Treatment)
levels(ARA_reduced$Treatment)[levels(ARA_reduced$Treatment) == "15N2_C2H2_CK"] <- "Stem gas"
levels(ARA_reduced$Treatment)[levels(ARA_reduced$Treatment) == "15N2_C2H2_Air"] <- "Air"
levels(ARA_reduced$Treatment)[levels(ARA_reduced$Treatment) == "15N2_CK_CT"] <- "endogenous control"
levels(ARA_reduced$Treatment)[levels(ARA_reduced$Treatment) == "15N2_C2H2_CK_microCT"] <- "Control"
levels(ARA_reduced$Treatment)[levels(ARA_reduced$Treatment) == "15N2_C2H2_CK_nobarkCT"]<- "no bark control"


```
## Visualise the data
- Summarise data: N fix rate (ng N/g/day)

```{r}
ARA_ngN_g_day_groupby_treatment_time <- ARA_reduced %>% group_by(Treatment, incubation_time_days) %>% 
  dplyr::summarise(
    mean_Nfix = mean(Nfix_rate_ngN_g_day), 
    min_Nfix = min(Nfix_rate_ngN_g_day),
    max_Nfix = max(Nfix_rate_ngN_g_day),
    se_Nfix = sd(Nfix_rate_ngN_g_day)/sqrt(n())
  )

print(ARA_ngN_g_day_groupby_treatment_time)
write.table(ARA_ngN_g_day_groupby_treatment_time, "ARA_ngN_g_day_groupby_treatment_time.txt", sep = "\t")

```


- Only treatments and negative control
```{r}
ARA_summary_treatments<- ARA_ngN_g_day_groupby_treatment_time %>% 
  filter(Treatment %in% c('Stem gas','Air', 'Control'))
print(ARA_summary_treatments)
write.table(ARA_summary_treatments, "ARA_summary_treatments.txt", sep = "\t")

```


##Plot the data

- Ethylene ppm

```{r}

ARA_groupby_ethylene_ppm <- ARA_reduced %>% group_by(Treatment, incubation_time_days) %>% 
  dplyr::summarise(
    mean_ppm = mean(Concentration_ppm), 
    min_ppm = min(Concentration_ppm),
    max_ppm = max(Concentration_ppm),
    se_ppm = sd(Concentration_ppm)/sqrt(n())
  )

print(ARA_groupby_ethylene_ppm)
write.table(ARA_groupby_ethylene_ppm, "ARA_groupby_ethylene_ppm.txt", sep = "\t")

```



## Hypothesis testing
- Remove the treatments that we are not interested in testing. We only leave the negative control (sterile):
```{r}
ARA_reduced_no_ct <- ARA_reduced[!(ARA_reduced$Treatment %in% c("endogenous control", "no bark control")), ]
print(ARA_reduced_no_ct)
str(ARA_reduced_no_ct)

```

```{r}
# Define color palette
treatment_colors <- c(
  "Air" = "#009E73",   # bluish green
  "Stem gas"      = "#D55E00",   # vermillion
  "Control"  = "#0072B2"    # blue
)

# Plot
p_C2H4_days <- ggplot(ARA_reduced_no_ct, aes(x = incubation_time_days, y = Concentration_ppm, fill = Treatment)) +
  geom_boxplot(position = position_dodge(width = 0.75),
               outlier.shape = NA,
               width = 0.6,
               alpha = 0.7,
               linewidth = 0.5,
               color = "black") +
  geom_point(aes(fill = Treatment),
             position = position_jitterdodge(jitter.width = 0, dodge.width = 0.75),
             size = 2,
             alpha = 0.8,
             shape = 21,
             color = "black",
             stroke = 0.5) +
  scale_fill_manual(values = treatment_colors, 
                    name = "Treatment",
                    breaks = c("Air", "Stem gas", "Control")) +
  
  labs(y = expression(C[2]~H[4]~"concentration (ppm)"),
       x = "Incubation time (days)") +
    scale_y_continuous(limits = c(0, 30),
    expand = expansion(mult = c(0.05, 0.1))
  ) +
  theme_bw(base_size = 9) +
  theme(
    text = element_text(family = "sans", color="black"),   # Use Arial font
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    panel.border = element_rect(color="black", fill=NA, linewidth = 0.5),
    axis.ticks = element_line(color = "black", linewidth = 0.4),
    axis.ticks.length = unit(2, "pt"),
    axis.text.x = element_text(color = "black", size = 9),
    axis.text.y = element_text(color = "black", size = 9),
    axis.title.y = element_text(color = "black", size = 11, margin = margin(r=5)),
    axis.title.x = element_text(color = "black", size = 11, margin = margin(t = 5)),
    legend.position = c(0.02, 0.98), 
    legend.justification = c(0, 1),
    legend.title = element_blank(),
    legend.text = element_text(color = "black", size = 9),
    legend.background = element_blank(),
    legend.spacing.y = unit(0.5, "mm"),
    legend.box.spacing = unit(1, "mm"),
    plot.margin = margin(5, 5, 5, 5, "mm"),
    legend.key.size = unit(5, "mm"),
  )

  ggsave("Nfix_C2H4_Fig1.png", 
         plot = p_C2H4_days, 
         width = 140, height = 110, units = "mm", dpi = 300)
  
ggsave("Nfix_C2H4_Fig1.pdf",
       plot = p_C2H4_days,
       width = 140, height = 110, units = "mm",
       device = cairo_pdf,
       dpi = 300)

  print(p_C2H4_days)


```


Plot ARA for publication:

```{r}

# Define color palette
treatment_colors <- c(
  "Air" = "#009E73",   # bluish green
  "Stem gas"      = "#D55E00",   # vermillion
  "Control"  = "#0072B2"    # blue
)

# Plot
p_ARA_days <- ggplot(ARA_reduced_no_ct, aes(x = incubation_time_days, y = Nfix_rate_ngN_g_day, fill = Treatment)) +
  geom_boxplot(position = position_dodge(width = 0.75),
               outlier.shape = NA,
               width = 0.6,
               alpha = 0.7,
               linewidth = 0.5,
               color = "black") +
  geom_point(aes(fill = Treatment),
             position = position_jitterdodge(jitter.width = 0, dodge.width = 0.75),
             size = 2,
             alpha = 0.8,
             shape = 21,
             color = "black",
             stroke = 0.5) +
  scale_fill_manual(values = treatment_colors, 
                    name = "Treatment",
                    breaks = c("Air", "Stem gas", "Control")) +
  
  # 1. Horizontal line segment (Bar)
  annotate(
    "segment",
    x = 2.61, xend = 3.11, 
    y = 65, yend = 65, 
    color = "black", 
    linewidth = 0.5
  ) +
  
  # 2. Vertical tick marks
  annotate(
    "segment",
    x = 2.61, xend = 2.61, 
    y = 63, yend = 63, 
    color = "black", 
    linewidth = 0.5
  ) +
  annotate(
    "segment",
    x = 3.11, xend = 3.11, 
    y = 63, yend = 63, 
    color = "black", 
    linewidth = 0.5
  ) +
  annotate(
    "text",
    x = (2.61+3.11)/2, y = 66, 
    label = "***", 
    color = "black", 
    size = 3.5 
  ) +
  labs(y = expression(N~fixation~rate~(ng~N~g^{-1}~day^{-1})),
       x = "Incubation time (days)") +
    scale_y_continuous(limits = c(0, 67),
    expand = expansion(mult = c(0.0001, 0.05))
  ) +
  theme_bw(base_size = 9) +
  theme(
    text = element_text(family = "sans", color="black"),   # Use Arial font
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    panel.border = element_rect(color="black", fill=NA, linewidth = 0.5),
    axis.ticks = element_line(color = "black", linewidth = 0.4),
    axis.ticks.length = unit(2, "pt"),
    axis.text.x = element_text(color = "black", size = 9),
    axis.text.y = element_text(color = "black", size = 9),
    axis.title.y = element_text(color = "black", size = 11, margin = margin(r=5)),
    axis.title.x = element_text(color = "black", size = 11, margin = margin(t = 5)),
    legend.position = c(0.02, 0.98), 
    legend.justification = c(0, 1),
    legend.title = element_blank(),
    legend.text = element_text(color = "black", size = 9),
    legend.background = element_blank(),
    legend.spacing.y = unit(0.5, "mm"),
    legend.box.spacing = unit(1, "mm"),
    plot.margin = margin(5, 5, 5, 5, "mm"),
    legend.key.size = unit(5, "mm"),
  )

ggsave("Nfix_ARA_Fig1_v2.png", 
         plot = p_ARA_days, 
         width = 140, height = 110, units = "mm", dpi = 300)
  
ggsave("Nfix_ARA_Fig1.pdf",
       plot = p_ARA_days,
       width = 140, height = 110, units = "mm",
       device = cairo_pdf,
       dpi = 300)

  print(p_ARA_days)


  

```

```{r}
plot_ARA_combined <- (p_C2H4_days / p_ARA_days)

  ggsave("plot_ARA_fig1.png", 
         plot = plot_ARA_combined, 
         width = 140, height = 210, units = "mm", dpi = 300)
  
ggsave("plot_ARA_fig1.pdf",
       plot = plot_ARA_combined,
       width = 140, height = 210, units = "mm",
       device = cairo_pdf,
       dpi = 300)
```

- 2-Way ANOVA with Treatment and Time as fixed factors
```{r}
anova_ARA<-aov(Nfix_rate_ngN_g_day~Treatment+incubation_time_days+Treatment*incubation_time_days, data=ARA_reduced_no_ct)
summary(anova_ARA)



```
  - Check for ANOVA assumptions (normality and homogeneity of variance)
```{r}
residuals_ARA<-residuals(anova_ARA)
qqnorm(residuals_ARA)
qqline(residuals_ARA, col = "red")
hist(residuals_ARA)
shapiro.test(residuals_ARA)
leveneTest(Nfix_rate_ngN_g_day~Treatment*incubation_time_days, data=ARA_reduced_no_ct)
```

- Post-hoc comparison using Tukey's test. Separating effects (time and treatment). Better option:
```{r}

emm_ARA_by_time <- emmeans(anova_ARA, ~ Treatment | incubation_time_days)
pw_comp_ARA_by_time <- pairs(emm_ARA_by_time, adjust = "tukey")
print(pw_comp_ARA_by_time)
write.table(pw_comp_ARA_by_time, "pairwise_results_ARA_by_time.txt", sep = "\t")

emm_ARA_by_treatment <- emmeans(anova_ARA, ~ incubation_time_days | Treatment)
pw_comp_ARA_by_treatment <- pairs(emm_ARA_by_treatment, adjust = "tukey")
print(pw_comp_ARA_by_treatment)
write.table(pw_comp_ARA_by_treatment, "pairwise_results_ARA_by_treatment.txt", sep = "\t")

```


# 15N2 method 
## Read the data:

```{r}

N15 <- read.csv("Data/15N_data.csv", stringsAsFactors = TRUE)
```

## How does time and treatment affect N fixation using 15N2 method

## Variables - Treatment, Time, N fixation (ng/g/day)

## View the data and structure the data

- Reduce the data set to the columns we are interested in:
```{r}
N15$incubation_time_days <- as.factor(N15$incubation_time_days)
N15_reduced <-N15 %>% dplyr::select(ID, av_d15N,Treatment,incubation_time_days, ngN_g_day)
```

- Change the name of variables and treatments and remove the rows we are not interested in (natural abundance)
```{r}

N15_reduced$Treatment <- factor(N15_reduced$Treatment)
levels(N15_reduced$Treatment)[levels(N15_reduced$Treatment) == "C2H2 + CK"] <- "C2H2 + Stem gas"
levels(N15_reduced$Treatment)[levels(N15_reduced$Treatment) == "Cocktail"] <- "Stem gas"
levels(N15_reduced$Treatment)[levels(N15_reduced$Treatment) == "Control + CK"]<- "Control"
levels(N15_reduced$Treatment)[levels(N15_reduced$Treatment) == "CK + CH4-3"]<- "Stem gas + CH4"
levels(N15_reduced$Treatment)[levels(N15_reduced$Treatment) == "DFM + CK"]<- "DFM"

```


## Visualisation of the data:
- Summarise the data (mean, min, max and SE)
```{r}
N15_BNF_summary <- N15_reduced %>% group_by(Treatment, incubation_time_days) %>% 
  dplyr::summarise(
    mean_Nfix = mean(ngN_g_day), 
    min_Nfix = min(ngN_g_day),
    max_Nfix = max(ngN_g_day),
    se_Nfix = sd(ngN_g_day)/sqrt(n())
  )
print(N15_BNF_summary)
write.table(N15_BNF_summary, "N15_BNF_summary.txt", sep = "\t")

N15_d15N_summary <- N15_reduced %>% group_by(Treatment, incubation_time_days) %>% 
  dplyr::summarise(
    mean_d15N = mean(av_d15N), 
    min_d15N = min(av_d15N),
    max_d15N = max(av_d15N),
    se_d15N = sd(av_d15N)/sqrt(n())
  )
print(N15_d15N_summary)
write.table(N15_d15N_summary, "N15_d15N_summary.txt", sep = "\t")

```

- Summarise only 4 days:

```{r}

N15_BNF_summary_4days<- N15_BNF_summary %>% 
  filter(incubation_time_days == '4')

```


Plot for publication Figure 3: 
```{r}

N15_reduced$Treatment <- factor(N15_reduced$Treatment, 
                                levels = c("C2H2 + Air", 
                                          "C2H2 + Stem gas", 
                                          "Air", 
                                          "Stem gas", 
                                          "Stem gas + CH4", 
                                          "DFM", 
                                          "Control"))


full_treatment_colors <- c(
  "C2H2 + Air"      = "#AA3377",  # Purple
  "C2H2 + Stem gas" = "#EE6677",  # Red/Pink
  "Air"             = "#4477AA",  # Blue
  "Stem gas"        = "#CCBB44",  # Yellow
  "Stem gas + CH4"  = "#009988",  # Teal (CVD-safe)
  "DFM"             = "#BBBBBB",  # Grey
  "Control"         = "#66CCEE"   # Cyan
)


full_treatment_colors_publication <- c(
  "C2H2 + Air"      = "#CC79A7",  # Magenta (Okabe–Ito)
  "C2H2 + Stem gas" = "#D55E00",  # Orange 
  "Air"             = "#0072B2",  # Blue (Okabe–Ito)
  "Stem gas"        = "#F0E442",  # Yellow (Okabe–Ito)
  "Stem gas + CH4"  = "#009E73",  # Green-teal (Okabe–Ito)
  "DFM"             = "#999999",  # Grey (Okabe–Ito)
  "Control"         = "#56B4E9"   # Sky blue (Okabe–Ito)
)



# Plot
p_15N <- ggplot(N15_reduced, aes(x = factor(incubation_time_days), y = ngN_g_day, fill = Treatment)) +
  geom_boxplot(position = position_dodge(width = 0.8),
               outlier.shape = NA,
               width = 0.6,
               alpha = 0.7,
               linewidth = 0.4,
               color = "black") +
  geom_point(aes(fill = Treatment),
             position = position_jitterdodge(jitter.width = 0, dodge.width = 0.8),
             size = 1.8,
             alpha = 0.8,
             shape = 21,
             color = "black",
             stroke = 0.4) +
  geom_vline(
    xintercept = c(1.5, 2.5), # Halfway between 1&2 and 2&4 (if 4 is the 3rd factor level)
    linetype = "dotted",
    color = "black",
    linewidth = 0.3
  ) +
  scale_fill_manual(values = full_treatment_colors_publication, 
                    name = "Treatment",
                    labels = c(expression(C[2]*H[2]~+~"Air"), expression(C[2]*H[2]~+~"Stem gas"), "Air", "Stem gas", expression("Stem gas"~+~CH[4]), "DFM", "Control"),
                    breaks = c("C2H2 + Air", "C2H2 + Stem gas", "Air", "Stem gas", "Stem gas + CH4", "DFM", "Control"),
                    expand = expansion(add = 0.5)) +
  
  labs(y = expression(N~fixation~rate~(ng~N~g^{-1}~day^{-1})),
       x = "Incubation time (days)") +
    scale_y_continuous(limits = c(0, 210),
    expand = expansion(mult = c(0.02, 0.05))
  ) +
  theme_bw(base_size = 9) +
  theme(
    text = element_text(family = "sans", color="black"),   # Use Arial font
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    panel.border = element_rect(color="black", fill=NA, linewidth = 0.5),
    axis.ticks = element_line(color = "black", linewidth = 0.4),
    axis.ticks.length = unit(2, "pt"),
    axis.text.x = element_text(color = "black", size = 9),
    axis.text.y = element_text(color = "black", size = 9),
    axis.title.y = element_text(color = "black", size = 11, margin = margin(r=5)),
    axis.title.x = element_text(color = "black", size = 11, margin = margin(t = 5)),
    legend.position = c(0.01, 0.99), 
    legend.justification = c(0, 1),
    legend.title = element_blank(),
    legend.text = element_text(color = "black", size = 8),
    legend.text.align = 0,
    legend.background = element_blank(),
    legend.spacing.y = unit(1, "mm"),
    legend.box.spacing = unit(1, "mm"),
    plot.margin = margin(5, 5, 5, 5, "mm"),
    legend.key.size = unit(4, "mm"),
  )

ggsave("Nfix_15N_Fig3_v3.png", 
         plot = p_15N, 
         width = 180, height = 130, units = "mm", dpi = 300)
  
ggsave("Nfix_15N_Fig3_v2.pdf",
       plot = p_15N,
       width = 180, height = 130, units = "mm",
       device = cairo_pdf,
       dpi = 300)

  print(p_15N)
  

  
```


## Hypothesis testing:

- ANOVA 2 ways: Gas Cocktail vs Air vs C2H2 vs DFM vs Controls 
```{r}

anova_model_N15 <- aov(ngN_g_day~ Treatment*incubation_time_days, data = N15_reduced)

summary(anova_model_N15)
```
  - ANOVA assumptions
```{r}

residuals_N15<-residuals(anova_model_N15)

qqnorm(residuals_N15)
qqline(residuals_N15, col = "red")
hist(residuals_N15)
shapiro.test(residuals_N15)

leveneTest(ngN_g_day~ Treatment*incubation_time_days , data = N15_reduced)


```
- Data not normal. Let's try to use the function Boxcox to determine the best transformation:
```{r}
lm_model <- lm(ngN_g_day ~ Treatment * incubation_time_days, data= N15_reduced)

Boxcox_Result <- boxcox(lm_model)

lambda_best <- Boxcox_Result$x[which.max(Boxcox_Result$y)]
print(lambda_best)
```

- Transformation of data and ANOVA 
```{r}

N15_reduced$boxcox_transformed <- (N15_reduced$ngN_g_day^lambda_best - 1) / lambda_best

anova_model_N15_transformed <- aov(boxcox_transformed~ Treatment*incubation_time_days, data = N15_reduced)

summary(anova_model_N15_transformed)
residuals_N15_transformed<-residuals(anova_model_N15_transformed)

qqnorm(residuals_N15_transformed)
qqline(residuals_N15_transformed, col = "red")
hist(residuals_N15_transformed)
shapiro.test(residuals_N15_transformed)

leveneTest(boxcox_transformed~ Treatment*incubation_time_days , data = N15_reduced)

emm_15N_aov_transformed_treatment <- emmeans(anova_model_N15_transformed, ~ Treatment|incubation_time_days)
pw_comp_15N_aov_transformed_treatment <- pairs(emm_15N_aov_transformed_treatment, adjust = "BH")
print(pw_comp_15N_aov_transformed_treatment)
write.table(pw_comp_15N_aov_transformed_treatment, "pairwise_results_15N_aov_transformed_treatment.txt", sep = "\t")

emm_15N_aov_transformed_time <- emmeans(anova_model_N15_transformed, ~ incubation_time_days|Treatment)
pw_comp_15N_aov_transformed_time <- pairs(emm_15N_aov_transformed_time, adjust = "BH")
print(pw_comp_15N_aov_transformed_time)
write.table(pw_comp_15N_aov_transformed_time, "pairwise_results_15N_aov_transformed_time.txt", sep = "\t")


```
- GLMs 
```{r}
glm_model_15N <- glm(ngN_g_day ~ Treatment * incubation_time_days, 
                     family = Gamma(link = "log"), 
                     data = N15_reduced)
summary(glm_model_15N)
car::Anova(glm_model_15N, type = "III", test.statistic = "F")

```
  - Check assumptions of GLM
```{r}
ggplot(data.frame(x1=N15_reduced$Treatment, pearson=residuals(glm_model_15N, type="pearson")),
       aes(x=N15_reduced$Treatment, y=pearson))+geom_point() +theme_bw()
plot(glm_model_15N)
shapiro.test(residuals(glm_model_15N))
leveneTest(glm_model_15N)

# DHARMa simulated residuals
sim_res_15N <- simulateResiduals(glm_model_15N)
plot(sim_res_15N)
testResiduals(sim_res_15N)

AIC(glm_model_15N)
AIC(anova_model_N15_transformed)


```

- Pairwise comparisons using Tukey's test
```{r}

# Pairwise comparisons for Treatment at each level of incubation_time_days
pw_comp_15Nglm_treatment<- emmeans(glm_model_15N, ~ Treatment | incubation_time_days)
print(pairs(pw_comp_15Nglm_treatment, adjust = "Tukey"))

write.table(pairs(pw_comp_15Nglm_treatment, adjust = "Tukey"),
            "pw_comp_15Nglm_treatment.txt", sep = "\t")

pw_comp_15Nglm_time <- emmeans(glm_model_15N, ~ incubation_time_days | Treatment)
print(pairs(pw_comp_15Nglm_time, adjust = "Tukey"))

write.table(pairs(pw_comp_15Nglm_time, adjust = "Tukey"),
            "pw_comp_15Nglm_time.txt", sep = "\t")
```



- Prediction values:
```{r}
# Add predicted values to your dataset
N15_reduced$predicted_values <- predict(glm_model_15N, type = "response")
summary(N15_reduced)

```

- Residual plot:

```{r}
# Calculate residuals
N15_reduced$residuals <- residuals(glm_model_15N, type = "deviance")

# Plot residuals
ggplot(N15_reduced, aes(x = predicted_values, y = residuals, color = Treatment)) +
  geom_point(size = 3) +
  geom_hline(yintercept = 0, linetype = "dashed", color = "red") +
  labs(title = "Residuals Plot",
       x = "Predicted Values",
       y = "Residuals") +
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5)) +
  scale_color_brewer(palette = "Set3")

```

The model is better with the GLM.


# ARA vs 15N2 method 

## Questions: 
- Differences in N fixation between treatments
- Differences in N fixation between methods used
- Differences in N fixation between incubation times

## View the data and structure the data
- Create a new column N fixation rates combining N fix rates from ARA and 15N method. Then add a column called method to be able to compare both methods. all=FALSE to remove the treatments that are not shared between method (NA)

```{r}
ARA_vs_15N_4days <- read.csv("Data/ARA_vs_15N_data.csv", stringsAsFactors = TRUE)
ARA_vs_15N_4days$incubation_time_days<-as.factor(ARA_vs_15N_4days$incubation_time_days)

```
- Rename the Methods values to ARA and 15N

- Select only the columns of interest

```{r}
write.table(ARA_vs_15N_4days, "ARA_vs_15N_4days.txt", sep = "\t")
```

## Data visualisation

- Summarise the data
```{r, message=FALSE}
#Version 1
ARA_vs_15N_4days_summary <- ARA_vs_15N_4days %>% group_by(Treatment, Method) %>% 
  dplyr::summarise(
    mean_Nfix = mean(N_Fixation_Rates_ngN_g_day), 
    min_Nfix = min(N_Fixation_Rates_ngN_g_day),
    max_Nfix = max(N_Fixation_Rates_ngN_g_day),
    se_Nfix = sd(N_Fixation_Rates_ngN_g_day)/sqrt(n())
  )
print(ARA_vs_15N_4days_summary)
write.table(ARA_vs_15N_4days_summary, "N_fix_rates_summary_ARA_vs_15N.txt", sep = "\t")

```

- Plot the data
```{r, echo=FALSE}

method_colors <- c(
  "15N2"      = "#009E73",
  "ARA" = "#D55E00")

# Plot
p_ARA_vs_15N <- ggplot(ARA_vs_15N_4days, aes(x = Treatment, y = N_Fixation_Rates_ngN_g_day, fill = Method)) +
  geom_boxplot(position = position_dodge(width = 0.75),
               outlier.shape = NA,
               width = 0.6,
               alpha = 0.7,
               linewidth = 0.5,
               color = "black") +
  geom_point(aes(fill = Method),
             position = position_jitterdodge(jitter.width = 0, dodge.width = 0.75),
             size = 2,
             alpha = 0.8,
             shape = 21,
             color = "black",
             stroke = 0.5) +
  scale_fill_manual(values = method_colors, 
                    name = "Method",
                    labels = c(expression({}^15*N[2]), "ARA"),
                    breaks = c("15N2", "ARA")) +
  scale_x_discrete(
    # Assuming your current Treatment labels in the data are "C2H2 + stem gas" and "C2H2 + air"
    labels = c(
      expression(C[2]*H[2]~+~"Air"), # C[2]H[2] for subscripts, use ~ for spacing
      expression(C[2]*H[2]~+~"Stem gas"), "Stem gas")) +
  labs(y = expression(N~fixation~rate~(ng~N~g^{-1}~day^{-1})),
       x = "Treatment") +
  annotate(
    "text",
    x = 2.81, y = 153, 
    label = "**", 
    color = "black", 
    size = 4 
  ) +
    scale_y_continuous(limits = c(0, 155),
    expand = expansion(mult = c(0.05, 0.1))
  ) +
  theme_bw(base_size = 9) +
  theme(
    text = element_text(family = "sans", color="black"),   # Use Arial font
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    panel.border = element_rect(color="black", fill=NA, linewidth = 0.5),
    axis.ticks = element_line(color = "black", linewidth = 0.4),
    axis.ticks.length = unit(2, "pt"),
    axis.text.x = element_text(color = "black", size = 9),
    axis.text.y = element_text(color = "black", size = 9),
    axis.title.y = element_text(color = "black", size = 11, margin = margin(r=5)),
    axis.title.x = element_text(color = "black", size = 11, margin = margin(t = 5)),
    legend.position = c(0.02, 0.98), 
    legend.justification = c(0, 1),
    legend.title = element_blank(),
    legend.text = element_text(color = "black", size = 9),
    legend.background = element_blank(),
    legend.spacing.y = unit(0.5, "mm"),
    legend.box.spacing = unit(1, "mm"),
    plot.margin = margin(5, 5, 5, 5, "mm"),
    legend.key.size = unit(5, "mm"),
  )

  ggsave("ARA_vs_15N_Fig2.png", 
         plot = p_ARA_vs_15N, 
         width = 140, height = 110, units = "mm", dpi = 300)
  
ggsave("ARA_vs_15N_Fig2.pdf",
       plot = p_ARA_vs_15N,
       width = 140, height = 110, units = "mm",
       device = cairo_pdf,
       dpi = 300)

  print(p_ARA_vs_15N)

```

## Stats

- 2-way ANOVA (only consider Method and Treatment after 4 days incubation to simplify)
```{r}
#Version 1
anova_ARAvs_15N<-aov(N_Fixation_Rates_ngN_g_day~Treatment+Method+Treatment*Method, data=ARA_vs_15N_4days)
summary(anova_ARAvs_15N)


```

- ANOVA assumptions

```{r}
#Version1
residuals_ARA_vs_15N<-residuals(anova_ARAvs_15N)
qqnorm(residuals_ARA_vs_15N)
qqline(residuals_ARA_vs_15N, col = "red")
hist(residuals_ARA_vs_15N)
shapiro.test(residuals_ARA_vs_15N)
leveneTest(N_Fixation_Rates_ngN_g_day~ Treatment, data = ARA_vs_15N_4days)

```


- GLM model using family gamma log
```{r}
#Version 1

glm_model_ARA_vs_15N_log <- glm(N_Fixation_Rates_ngN_g_day ~ Treatment*Method, 
                     family = Gamma(link = "log"), 
                     data = ARA_vs_15N_4days)
summary(glm_model_ARA_vs_15N_log)

car::Anova(glm_model_ARA_vs_15N_log, type = "III", test.statistic = "F")

```

  - Assumptions of the GLM model
```{r}
#version1
ggplot(data.frame(x1=ARA_vs_15N_4days$Treatment, pearson=residuals(glm_model_ARA_vs_15N_log, type="pearson")),
       aes(x=x1, y=pearson))+geom_point() +theme_bw()
plot(glm_model_ARA_vs_15N_log)

shapiro.test(residuals(glm_model_ARA_vs_15N_log))
leveneTest(glm_model_ARA_vs_15N_log)


# DHARMa simulated residuals
sim_res_ARA_vs_15N <- simulateResiduals(glm_model_ARA_vs_15N_log)
plot(sim_res_ARA_vs_15N)
testResiduals(sim_res_ARA_vs_15N)
```


Skewed Distribution: Nitrogen fixation rates often follow a right-skewed distribution (many small values and a few large values), and the log link is well-suited for handling this kind of data by modeling the exponential relationship between predictors and the response variable.

Interpretability: The log link provides more straightforward interpretation, as the coefficients represent percentage changes in nitrogen fixation rates, which is intuitive in ecological studies where rates often scale multiplicatively (e.g., "a doubling of the rate in response to this factor").



- Pairwise comparisons using two methods (Version 1-considering stem gas as endogenous control)
```{r}
#Version 1

# Pairwise comparisons for Treatment 
pw_comp_15N_ARA_treatment<- emmeans(glm_model_ARA_vs_15N_log, ~ Treatment | Method)
print(pairs(pw_comp_15N_ARA_treatment, adjust = "tukey"))

write.table(pairs(pw_comp_15N_ARA_treatment, adjust = "tukey"),
            "pw_comp_15N_ARA_treatment.txt", sep = "\t")

# Pairwise comparisons for method 
pw_comp_15N_ARA_method <- emmeans(glm_model_ARA_vs_15N_log, ~ Method | Treatment)
print(pairs(pw_comp_15N_ARA_method, adjust = "tukey"))

write.table(pairs(pw_comp_15N_ARA_method, adjust = "tukey"),
            "pw_comp_15N_ARA_method.txt", sep = "\t")

# Pairwise comparisons for method 
pw_comp_15N_ARA_method_treatment <- emmeans(glm_model_ARA_vs_15N_log, ~ Method * Treatment)
print(pairs(pw_comp_15N_ARA_method_treatment, adjust = "tukey"))

write.table(pairs(pw_comp_15N_ARA_method_treatment, adjust = "tukey"),
            "pw_comp_15N_ARA_method.txt", sep = "\t")
```


