---
title: "Pilot_CH4_paperbark"
author: "Paula Gomez - SCU"
date: "`r Sys.Date()`"
output:
  word_document: default
  keep_tex: false
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r cars, echo=FALSE, warning=FALSE, message=FALSE}
library(tidyverse)
library(lme4)
library(Matrix)
library(dplyr)
library(rstatix)
library(lmerTest)
library(car)
library(emmeans)
library(nlme)
library(lme4)
library(stats)
library(MASS)
library(performance)
library(DHARMa)
library(ggplot2)
library(ggpubr)

```

#Read the data

```{r}

bark_data_2025 <- read.csv('Data/data_CH4.csv', stringsAsFactors = TRUE)
```

# Research Questions:

-   Differences between treatments
-   Differences between time steps

# View the data and structure the data:

- Clean up the data:

```{r}
bark_data_2025$incubation_time<-as.factor(bark_data_2025$incubation_time)

str(bark_data_2025)

```

- We have two reading of the IRMS per sample, so we make an average for d15N and average for N fix rates (ngN/g/day):

```{r, message=FALSE}
av_15N_bark_data_2025 <- bark_data_2025 %>% 
  group_by(ID, treatment, incubation_time) %>% 
  summarise(across(d15N,mean))

av_Nfix2_bark_data_2025 <- bark_data_2025 %>% 
  group_by(ID, treatment, incubation_time) %>% 
  summarise(across(ngN_g_day_2,mean))
```

- Summarise the data

```{r}

summary_Nfix2_bark_data_2025 <- av_Nfix2_bark_data_2025 %>% group_by(treatment, incubation_time) %>% 
  dplyr::summarise(
    mean_Nfix = mean(ngN_g_day_2), 
    min_Nfix = min(ngN_g_day_2),
    max_Nfix = max(ngN_g_day_2),
    se_Nfix = sd(ngN_g_day_2)/sqrt(n())
  )
print(summary_Nfix2_bark_data_2025)
write.table(summary_Nfix2_bark_data_2025, "summary_Nfix2_bark_data_2025.txt", sep = "\t")

summary_d15N_bark_data_2025 <- av_15N_bark_data_2025 %>% group_by(treatment, incubation_time) %>% 
  dplyr::summarise(
    mean_d15N = mean(d15N), 
    min_d15N = min(d15N),
    max_d15N = max(d15N),
    se_d15N = sd(d15N)/sqrt(n())
  )
print(summary_d15N_bark_data_2025)
write.table(summary_d15N_bark_data_2025, "summary_d15N_bark_data_2025.txt", sep = "\t")

```


Publication Fig 4

```{r}
av_15N_bark_data_2025$treatment <- factor(av_15N_bark_data_2025$treatment)

levels(av_15N_bark_data_2025$treatment)[levels(av_15N_bark_data_2025$treatment) == "15N2_CH4_O2"] <- 
"CH4 + O2"
levels(av_15N_bark_data_2025$treatment)[levels(av_15N_bark_data_2025$treatment) == "15N2"] <- 
"O2"
av_15N_bark_data_2025$treatment <- factor(av_15N_bark_data_2025$treatment, 
                                levels = c("O2", 
                                          "CH4 + O2", 
                                          "Control"))

treatment_colors <- c(
  "O2" = "#009E73",
  "CH4 + O2" = "#D55E00",
  "Control" =  "#0072B2")


p_d15N_CH4 <- ggplot(av_15N_bark_data_2025, aes(x = factor(incubation_time), y = d15N, fill = treatment)) +
  geom_boxplot(position = position_dodge(width = 0.8),
               outlier.shape = NA,
               width = 0.6,
               alpha = 0.7,
               linewidth = 0.4,
               color = "black") +
  geom_point(aes(fill = treatment),
             position = position_jitterdodge(jitter.width = 0, dodge.width = 0.8),
             size = 1.8,
             alpha = 0.8,
             shape = 21,
             color = "black",
             stroke = 0.4) +
  geom_vline(
    xintercept = c(1.5, 2.5, 3.5, 4.5), # Halfway between 1&2 and 2&4 (if 4 is the 3rd factor level)
    linetype = "dotted",
    color = "black",
    linewidth = 0.3
  ) +
  scale_fill_manual(values = treatment_colors, 
                    name = "Treatment",
                    labels = c(expression(O[2]), expression(CH[4]~+~O[2]), "Control"),
                    breaks = c("O2", "CH4 + O2", "Control"),
                    expand = expansion(add = 0.5)) +
  
  labs(y = expression(delta^{15}*N~("\u2030")),
       x = "Incubation time (days)") +
    scale_y_continuous(limits = c(0, 150),
    expand = expansion(mult = c(0.02, 0.05))
  ) +
  theme_bw(base_size = 9) +
  theme(
    text = element_text(family = "sans", color="black"),   # Use Arial font
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    panel.border = element_rect(color="black", fill=NA, linewidth = 0.5),
    axis.ticks = element_line(color = "black", linewidth = 0.4),
    axis.ticks.length = unit(2, "pt"),
    axis.text.x = element_text(color = "black", size = 9),
    axis.text.y = element_text(color = "black", size = 9),
    axis.title.y = element_text(color = "black", size = 11, margin = margin(r=5)),
    axis.title.x = element_text(color = "black", size = 11, margin = margin(t = 5)),
    legend.position = c(0.01, 0.99), 
    legend.justification = c(0, 1),
    legend.title = element_blank(),
    legend.text = element_text(color = "black", size = 8),
    legend.text.align = 0,
    legend.background = element_blank(),
    legend.spacing.y = unit(1, "mm"),
    legend.box.spacing = unit(1, "mm"),
    plot.margin = margin(5, 5, 5, 5, "mm"),
    legend.key.size = unit(4, "mm"),
  )

  ggsave("d15N_CH4.png", 
         plot = p_d15N_CH4, 
         width = 180, height = 130, units = "mm", dpi = 300)
  
ggsave("d15N_CH4.pdf",
       plot = p_d15N_CH4,
       width = 180, height = 130, units = "mm",
       device = cairo_pdf,
       dpi = 300)

  print(p_d15N_CH4)

```



```{r}
av_Nfix2_bark_data_2025$treatment <- factor(av_Nfix2_bark_data_2025$treatment)
levels(av_Nfix2_bark_data_2025$treatment)[levels(av_Nfix2_bark_data_2025$treatment) == "15N2_CH4_O2"] <- 
"CH4 + O2"
levels(av_Nfix2_bark_data_2025$treatment)[levels(av_Nfix2_bark_data_2025$treatment) == "15N2"] <- 
"O2"
av_Nfix2_bark_data_2025$treatment <- factor(av_Nfix2_bark_data_2025$treatment, 
                                levels = c("O2", 
                                          "CH4 + O2", 
                                          "Control"))
# Define color palette
treatment_colors <- c(
  "O2" = "#009E73",
  "CH4 + O2" = "#D55E00",
  "Control" =  "#0072B2")

# Plot
p_Nfix_CH4 <- ggplot(av_Nfix2_bark_data_2025, aes(x = factor(incubation_time), y = ngN_g_day_2, fill = treatment)) +
  geom_boxplot(position = position_dodge(width = 0.8),
               outlier.shape = NA,
               width = 0.6,
               alpha = 0.7,
               linewidth = 0.4,
               color = "black") +
  geom_point(aes(fill = treatment),
             position = position_jitterdodge(jitter.width = 0, dodge.width = 0.8),
             size = 1.8,
             alpha = 0.8,
             shape = 21,
             color = "black",
             stroke = 0.4) +
  geom_vline(
    xintercept = c(1.5, 2.5, 3.5, 4.5), # Halfway between 1&2 and 2&4 (if 4 is the 3rd factor level)
    linetype = "dotted",
    color = "black",
    linewidth = 0.3
  ) +
  scale_fill_manual(values = treatment_colors, 
                    name = "Treatment",
                    labels = c(expression(O[2]), expression(CH[4]~+~O[2]), "Control"),
                    breaks = c("O2", "CH4 + O2", "Control"),
                    expand = expansion(add = 0.5)) +
  
  labs(y = expression(N~fixation~rate~(ng~N~g^{-1}~day^{-1})),
       x = "Incubation time (days)") +
    scale_y_continuous(limits = c(0, 350),
    expand = expansion(mult = c(0.02, 0.05))
  ) +
  theme_bw(base_size = 9) +
  theme(
    text = element_text(family = "sans", color="black"),   # Use Arial font
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    panel.border = element_rect(color="black", fill=NA, linewidth = 0.5),
    axis.ticks = element_line(color = "black", linewidth = 0.4),
    axis.ticks.length = unit(2, "pt"),
    axis.text.x = element_text(color = "black", size = 9),
    axis.text.y = element_text(color = "black", size = 9),
    axis.title.y = element_text(color = "black", size = 11, margin = margin(r=5)),
    axis.title.x = element_text(color = "black", size = 11, margin = margin(t = 5)),
    legend.position = c(0.01, 0.99), 
    legend.justification = c(0, 1),
    legend.title = element_blank(),
    legend.text = element_text(color = "black", size = 8),
    legend.text.align = 0,
    legend.background = element_blank(),
    legend.spacing.y = unit(1, "mm"),
    legend.box.spacing = unit(1, "mm"),
    plot.margin = margin(5, 5, 5, 5, "mm"),
    legend.key.size = unit(4, "mm"),
  )

  ggsave("Nfix_CH4.png", 
         plot = p_Nfix_CH4, 
         width = 180, height = 130, units = "mm", dpi = 300)
  
ggsave("Nfix_CH4.pdf",
       plot = p_Nfix_CH4,
       width = 180, height = 130, units = "mm",
       device = cairo_pdf,
       dpi = 300)

  print(p_Nfix_CH4)
```

Plot CH4 with O2: 
```{r}
o2_data <- read.csv('Data/data_O2_CH4.csv', stringsAsFactors = TRUE)

Nfix_O2_data_2025 <- merge(av_Nfix2_bark_data_2025, 
                                  o2_data[, c("ID", "O2_.")], 
                                  by = "ID", 
                                  all.x = TRUE)


Nfix_O2_data_2025 <- Nfix_O2_data_2025[Nfix_O2_data_2025$treatment != "Control", ]

Nfix_O2_data_2025$treatment <- factor(Nfix_O2_data_2025$treatment, 
                                             levels = c("O2", "CH4 + O2"))


treatment_colors <- c(
  "O2" = "#009E73",
  "CH4 + O2" = "#D55E00")


oxygen_summary <- Nfix_O2_data_2025 %>%
  group_by(treatment, incubation_time) %>%
  summarise(
    mean_oxygen = mean(O2_., na.rm = TRUE),
    se_oxygen = sd(O2_., na.rm = TRUE) / sqrt(n()),
    .groups = 'drop'
  )


scaling_factor <- 350/20

# Plot with secondary axis for oxygen
p_Nfix_vs_O2_CH4 <- ggplot(Nfix_O2_data_2025, aes(x = factor(incubation_time), y = ngN_g_day_2, fill = treatment)) +
   # Add horizontal lines for O2 reference levels
  geom_hline(yintercept = 2 * scaling_factor, 
             linetype = "dashed", 
             color = "gray50", 
             linewidth = 0.5) +
  geom_hline(yintercept = 13.8 * scaling_factor, 
             linetype = "dashed", 
             color = "gray50", 
             linewidth = 0.5) +
  geom_boxplot(position = position_dodge(width = 0.8),
               outlier.shape = NA,
               width = 0.6,
               alpha = 0.8,
               linewidth = 0.4,
               color = "black") +
  geom_point(aes(fill = treatment),
             position = position_jitterdodge(jitter.width = 0, dodge.width = 0.8),
             size = 1.8,
             alpha = 0.8,
             shape = 21,
             color = "black",
             stroke = 0.4) +
  geom_vline(
    xintercept = c(1.5, 2.5, 3.5, 4.5),
    linetype = "dotted",
    color = "black",
    linewidth = 0.3
  ) +
  geom_line(data = oxygen_summary,
            aes(y = mean_oxygen * scaling_factor,
                x = factor(incubation_time),
                color = treatment,
                group = treatment),
            linewidth = 0.8,
            inherit.aes = FALSE) +
  # Add SE error bars
  geom_errorbar(data = oxygen_summary,
                aes(y = mean_oxygen * scaling_factor,
                    ymin = (mean_oxygen - se_oxygen) * scaling_factor,
                    ymax = (mean_oxygen + se_oxygen) * scaling_factor,
                    x = factor(incubation_time),
                    color = treatment),
                width = 0.15,
                linewidth = 0.4,
                color = "black",
                inherit.aes = FALSE) +
  # Add oxygen mean points
  geom_point(data = oxygen_summary,
             aes(y = mean_oxygen * scaling_factor,
                 x = factor(incubation_time),
                 fill = treatment),
             size = 2.5,
             shape = 23,
             color = "black",
             stroke = 0.6,
             inherit.aes = FALSE) +
     # Add text labels for O2 reference lines
  annotate("text", x = 2.7, y = 1 * scaling_factor, 
           label = expression(O[2]~inner~bark), 
           hjust = 0, vjust = -0.5, size = 3, color = "gray20") +
  annotate("text", x = 3.7, y = 13.8 * scaling_factor, 
           label = expression(O[2]~middle~bark), 
           hjust = 0, vjust = -0.5, size = 3, color = "gray20") +
  scale_fill_manual(values = treatment_colors, 
                    name = "Treatment",
                    labels = c(expression(O[2]), expression(CH[4]~+~O[2])),
                    breaks = c("O2", "CH4 + O2")) +
  scale_color_manual(values = treatment_colors,
                     name = expression(O[2]~concentration),
                     labels = c(expression(O[2]), expression(CH[4]~+~O[2])),
                     breaks = c("O2", "CH4 + O2")) +
  scale_y_continuous(
    limits = c(0, 350),
    expand = expansion(mult = c(0.02, 0.05)),
    sec.axis = sec_axis(~./scaling_factor, name = expression(O[2]~concentration~("%")),
                        breaks = seq(0, 20, by =5))
  ) +
  labs(y = expression(N~fixation~rate~(ng~N~g^{-1}~day^{-1})),
       x = "Incubation time (days)") +
  guides(
    fill = guide_legend(order = 1, 
                        override.aes = list(shape = 21, size = 1.8, alpha = 0.8)),
    color = guide_legend(order = 2,
                         override.aes = list(linetype = 1, linewidth = 0.8, shape = 23, size = 3, stroke = 0.6, fill = treatment_colors, color=treatment_colors))
  ) +
  theme_bw(base_size = 9) +
  theme(
    text = element_text(family = "sans", color="black"),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    panel.border = element_rect(color="black", fill=NA, linewidth = 0.5),
    axis.ticks = element_line(color = "black", linewidth = 0.4),
    axis.ticks.length = unit(2, "pt"),
    axis.text.x = element_text(color = "black", size = 9),
    axis.text.y = element_text(color = "black", size = 9),
    axis.text.y.right = element_text(color = "black", size = 9),
    axis.title.y = element_text(color = "black", size = 11, margin = margin(r=5)),
    axis.title.y.right = element_text(color = "black", size = 11, margin = margin(l=5)),
    axis.title.x = element_text(color = "black", size = 11, margin = margin(t = 5)),
    legend.position = c(0.01, 0.99), 
    legend.justification = c(0, 1),
    legend.title = element_blank(),
    legend.text = element_text(color = "black", size = 10),
    legend.text.align = 0,
    legend.background = element_blank(),
    legend.spacing.y = unit(0.5, "mm"),
    legend.box.spacing = unit(0.5, "mm"),
    plot.margin = margin(5, 5, 5, 5, "mm"),
    legend.key.size = unit(4, "mm"),
    legend.key = element_rect(fill = "white", color = NA)
  )

ggsave("Nfix_vs_O2_CH4_v2.png", 
       plot = p_Nfix_vs_O2_CH4, 
       width = 180, height = 130, units = "mm", dpi = 300)

ggsave("Nfix_vs_O2_CH4_v2.pdf",
       plot = p_Nfix_vs_O2_CH4,
       width = 180, height = 130, units = "mm",
       device = cairo_pdf,
       dpi = 300)

print(p_Nfix_vs_O2_CH4)
```


- Anova
```{r}

av_Nfix2_no_control <- subset(av_Nfix2_bark_data_2025, treatment != "Control")

av_15N_no_control <- subset(av_15N_bark_data_2025, treatment != "Control")
#anova BNF rate
anova_bark_2025_2<-aov(ngN_g_day_2~treatment+incubation_time+treatment*incubation_time, data=av_Nfix2_no_control)
summary(anova_bark_2025_2)
#anova d15N
anova_d15N_2025<-aov(d15N~treatment+incubation_time+treatment*incubation_time, data=av_15N_no_control)
summary(anova_d15N_2025)
```
- Assumptions
```{r}
#Assumptions BNF
residuals_bark_2025_2<-residuals(anova_bark_2025_2)
qqnorm(residuals_bark_2025_2)
qqline(residuals_bark_2025_2, col = "red")
hist(residuals_bark_2025_2)
shapiro.test(residuals_bark_2025_2)
leveneTest(ngN_g_day_2~treatment*incubation_time, data=av_Nfix2_no_control)

#Assumptions d15N
residuals_d15N_2025<-residuals(anova_d15N_2025)
qqnorm(residuals_d15N_2025)
qqline(residuals_d15N_2025, col = "red")
hist(residuals_d15N_2025)
shapiro.test(residuals_d15N_2025)
leveneTest(d15N~treatment*incubation_time, data=av_15N_no_control)
```


- Let's try with an GLM 

```{r}
#GLM for BNF

glm_model_no_control <- glm(ngN_g_day_2 ~ treatment * incubation_time,
                             family = Gamma(link = "log"),
                             data = av_Nfix2_no_control)
car::Anova(glm_model_no_control, type = "III", test.statistic = "F")


```

-   Check assumptions of GLM

```{r}

#Assumptions BNF no control

sim_res_no_control <- simulateResiduals(glm_model_no_control)
plot(sim_res_no_control)


ggplot(data.frame(x1=av_Nfix2_no_control$treatment, pearson=residuals(glm_model_no_control, type="pearson")),
       aes(x=av_Nfix2_no_control$treatment, y=pearson))+geom_point() +theme_bw()
plot(glm_model_no_control)

# Deviance residuals
plot(glm_model_no_control$fitted.values, residuals(glm_model_no_control, type = "deviance"),
     xlab = "Fitted Values", ylab = "Deviance Residuals")
abline(h = 0, lty = 2)

# Pearson residuals
plot(glm_model_no_control$fitted.values, residuals(glm_model_no_control, type = "pearson"),
     xlab = "Fitted Values", ylab = "Pearson Residuals")
abline(h = 0, lty = 2)


```

## Pairwise comparisons:

```{r}
# BNF for treatment comparisons at each time point no control
emm.trt.time_no_control <- emmeans(glm_model_no_control, ~ treatment | incubation_time)
pairs.trt.by.time_no_control <- pairs(emm.trt.time_no_control, adjust = "Tukey")
summary_trt_by_time_no_control <- summary(pairs.trt.by.time_no_control, type = "response")

# BNF for time comparisons within each treatment
emm.time.by.trt_no_control <- emmeans(glm_model_no_control, ~ incubation_time | treatment)
pairs.time.by.trt_no_control <- pairs(emm.time.by.trt_no_control, adjust = "Tukey")
summary_time_by_trt_no_control <- summary(pairs.time.by.trt_no_control, type = "response")

print("Pairwise Comparisons between Treatments at Each Incubation Time:")
print(summary_trt_by_time_no_control)
write.table(summary_trt_by_time_no_control, "summary_trt_by_time_no_control.txt", sep = "\t")

print("\nPairwise Comparisons between Incubation Times within Each Treatment:")
print(summary_time_by_trt_no_control)
write.table(summary_time_by_trt_no_control, "summary_time_by_trt_no_control.txt", sep = "\t")

```
```{r}
# Restrict to the days where Control has data
av_Nfix2_day3_7 <- subset(av_Nfix2_bark_data_2025, incubation_time %in% c(3, 7))
av_Nfix2_day3_7$incubation_time <- droplevels(factor(av_Nfix2_day3_7$incubation_time))

glm_model_day3_7 <- glm(ngN_g_day_2 ~ treatment * incubation_time,
                         family = Gamma(link = "log"),
                         data = av_Nfix2_day3_7)


car::Anova(glm_model_day3_7, type = "III", test.statistic = "F")

library(emmeans)
pairs(emmeans(glm_model_day3_7, ~ treatment | incubation_time))   # O2 vs CH4+O2 vs Control, at day 3 and day 7
pairs(emmeans(glm_model_day3_7, ~ incubation_time | treatment))   # each treatment's day3-vs-day7, including Control's now
pairs(emmeans(glm_model_day3_7, ~ treatment | incubation_time, type = "response"))
pairs(emmeans(glm_model_day3_7, ~ incubation_time | treatment, type = "response"))

```

# Oxygen data
#Read the data

```{r}

d15N_O2_data_2025 <- merge(av_15N_bark_data_2025, 
                                  o2_data[, c("ID", "O2_.")], 
                                  by = "ID", 
                                  all.x = TRUE)

# Filter out Control treatment
d15N_O2_data_2025  <- d15N_O2_data_2025[d15N_O2_data_2025 $treatment != "Control", ]

d15N_O2_data_2025$treatment <- factor(d15N_O2_data_2025$treatment)
levels(d15N_O2_data_2025$treatment)[levels(d15N_O2_data_2025$treatment) == "15N2_CH4_O2"] <- 
"CH4 + O2"
levels(d15N_O2_data_2025$treatment)[levels(d15N_O2_data_2025$treatment) == "15N2"] <- 
"O2"

d15N_O2_data_2025$treatment <- factor(d15N_O2_data_2025$treatment, 
                                             levels = c("O2", "CH4 + O2"))

```
- Summarise the data
```{r}
summary_O2_data <- o2_data %>% group_by(treatment, incubation_time) %>% 
  dplyr::summarise(
    mean_O2 = mean(O2_.), 
    min_O2 = min(O2_.),
    max_O2 = max(O2_.),
    se_O2 = sd(O2_.)/sqrt(n())
  )
print(summary_O2_data)
write.table(summary_O2_data, "summary_O2_data.txt", sep = "\t")
```

## Anova and assumptions
```{r}

o2_data_no_control  <- d15N_O2_data_2025[d15N_O2_data_2025 $treatment != "Control", ]
anova_O2<-aov(O2_.~treatment+incubation_time+treatment*incubation_time, data=o2_data_no_control)
summary(anova_O2)

residuals_O2<-residuals(anova_O2)
qqnorm(residuals_O2)
qqline(residuals_O2, col = "red")
hist(residuals_O2)
shapiro.test(residuals_O2)
leveneTest(O2_.~treatment, data=o2_data_no_control)
```

- Pairwise comparisons
```{r}

# treatment comparisons at each time point
emm_O2_trt_time <- emmeans(anova_O2, ~ treatment | incubation_time)
pairs_O2_trt_by_time <- pairs(emm_O2_trt_time, adjust = "Tukey")
summary_O2_trt_by_time <- summary(pairs_O2_trt_by_time, type = "response")

# comparisons between incubation times within each treatment
emm_O2_time_trt <- emmeans(anova_O2, ~ incubation_time | treatment)

pairs_O2_time_by_trt <- pairs(emm_O2_time_trt, adjust = "Tukey")
summary_O2_time_by_trt <- summary(pairs_O2_time_by_trt, type = "response")

print("O2 Pairwise Comparisons between Treatments at Each Incubation Time:")
print(summary_O2_trt_by_time)
write.table(summary_O2_trt_by_time, "summary_O2_trt_by_time.txt", sep = "\t")

print("\n O2 Pairwise Comparisons between Incubation Times within Each Treatment:")
print(summary_O2_time_by_trt)
write.table(summary_O2_time_by_trt, "summary_O2_time_by_trt.txt", sep = "\t")
```

- Calculate correlation BNF and O2

```{r}

ggscatter(Nfix_O2_data_2025, x = "O2_.", y = "ngN_g_day_2", 
          add = "reg.line", conf.int = TRUE, 
          cor.coef = TRUE, cor.method = "pearson",
          xlab = "O2 %", ylab = "BNF rate (ng N/g/day)")

shapiro.test(Nfix_O2_data_2025$O2_.)
shapiro.test(Nfix_O2_data_2025$ngN_g_day_2)

ggqqplot(Nfix_O2_data_2025$O2_., ylab = "O2 %")

ggqqplot(Nfix_O2_data_2025$ngN_g_day_2, ylab = "BNF rate")

#Data is not normal so we use spearman method:

correlation <- cor.test(Nfix_O2_data_2025$ngN_g_day_2, Nfix_O2_data_2025$O2_., method = "spearman")
correlation

#Separating correlations by treatment:

correlations_by_treatment <- Nfix_O2_data_2025 %>%
  group_by(treatment) %>%
  group_map(~ broom::tidy(cor.test(.x$O2_., .x$ngN_g_day_2, method = "spearman")) %>%
              mutate(treatment = unique(.x$treatment)),
            .keep = TRUE) %>%
  bind_rows()

cor_annotation <- correlations_by_treatment %>%
  mutate(
    label = paste0("rho = ", round(estimate, 2), "\n", "p = ", format.pval(p.value, digits = 2)),
    x = 5,  # or adjust as needed
    y = max(Nfix_O2_data_2025$ngN_g_day_2, na.rm = TRUE) * 0.95
  )
cor_annotation
write.table(cor_annotation, "cor_annotation_BNF_O2.txt", sep = "\t")
```
- Plot

```{r}
scatter_plot_BNF_O2 <- ggplot(Nfix_O2_data_2025, aes(x = O2_., y = ngN_g_day_2, color = treatment)) +
  geom_point(size = 3, alpha = 0.7) + 
  labs(
    x = "Oxygen Concentration (%)",
    y = "Nitrogen Fixation Rate (ng N/g/day)") +
  theme_bw() +
  theme(
    plot.title = element_text(hjust = 0.5), # Center the title
    axis.title.x = element_text(face = "bold", size = 12),
    axis.title.y = element_text(face = "bold", size = 12),
    axis.text.x = element_text(size = 10),
    axis.text.y = element_text(size = 10),
  panel.grid.major = element_blank(),
  panel.grid.minor = element_blank()
  ) +
  annotate(
    "text",
    x = min(Nfix_O2_data_2025$O2_.), # Adjust x position as needed
    y = max(Nfix_O2_data_2025$ngN_g_day_2), # Adjust y position as needed
    hjust = 0, # Left-justify text
    vjust = 1, # Top-justify text
    label = paste("Spearman's rho =", round(correlation$estimate, 2),
                  "\np-value =", format(correlation$p.value, digits = 3)),
    size = 4,
    color = "darkred" 
  )

print(scatter_plot_BNF_O2)


ggsave(filename = "scatter_plot_BNF_O2.png",
       plot = scatter_plot_BNF_O2,
       width = 8,
       height = 6,
       units = "in",
       dpi = 600) 
```

- No correlation with N fixation. However, I could check correlation with d15N: 
```{r}
ggscatter(d15N_O2_data_2025, x = "O2_.", y = "d15N", 
          add = "reg.line", conf.int = TRUE, 
          cor.coef = TRUE, cor.method = "pearson",
          xlab = "O2 %", ylab = "d15N")

shapiro.test(d15N_O2_data_2025$O2_.)
shapiro.test(d15N_O2_data_2025$d15N)

ggqqplot(d15N_O2_data_2025$O2_., ylab = "O2 %")

ggqqplot(d15N_O2_data_2025$d15N, ylab = "d15N")

#Data is not normal so we use spearman method:

cor_all <- cor.test(d15N_O2_data_2025$O2_., d15N_O2_data_2025$d15N, 
                    method = "spearman", exact = FALSE)

cor_O2 <- cor.test(d15N_O2_data_2025$O2_.[d15N_O2_data_2025$treatment == "O2"], 
                   d15N_O2_data_2025$d15N[d15N_O2_data_2025$treatment == "O2"], 
                   method = "spearman", exact = FALSE)

cor_CH4_O2 <- cor.test(d15N_O2_data_2025$O2_.[d15N_O2_data_2025$treatment == "CH4 + O2"], 
                       d15N_O2_data_2025$d15N[d15N_O2_data_2025$treatment == "CH4 + O2"], 
                       method = "spearman", exact = FALSE)
cor_all
cor_O2
cor_CH4_O2

```
- Plot

```{r}
legend_line <- data.frame(
  O2_. = c(0, 20),  
  d15N = c(0, 120),
  treatment = "All data",  
  legend_group = "All data"
)

p_correlation_d15N_O2 <- ggplot(d15N_O2_data_2025, aes(x = O2_., y = d15N, color = treatment, fill = treatment)) +
  geom_line(data = legend_line, aes(x=O2_., y = d15N, linetype = legend_group), color = "black", linewidth = 1.2, inherit.aes = FALSE)+
  # Add regression lines for each treatment
  geom_smooth(method = "lm", se = FALSE, linewidth = 1, alpha = 0.8) +
  # Add regression line for all data (black dashed)
  geom_smooth(aes(x = O2_., y = d15N), 
              method = "lm", se = FALSE, 
              color = "black", linetype = "dashed", linewidth = 1.2,
              inherit.aes = FALSE) +
  # Add points
  geom_point(size = 3.5, alpha = 0.8, shape = 21, color = "black", stroke = 0.5) +
  # Add correlation text for all data (top left)
  annotate("text", x = 0.5,
           y = 120,
           label = sprintf("Spearman's rho = %.2f\np-value = %.6f", 
                          cor_all$estimate, cor_all$p.value),
           hjust = 0, vjust = 1, size = 3.5, fontface = "bold") +
  # Add correlation text for CH4 + O2 (blue-gray)
  annotate("text", x = 0.5, 
           y = 90,
           label = sprintf("rho = %.2f\np = %.5f", 
                          cor_CH4_O2$estimate, cor_CH4_O2$p.value),
           hjust = 0, vjust = 1, size = 3.5, fontface = "bold", color = treatment_colors["CH4 + O2"]) +
  # Add correlation text for O2 (green)
  annotate("text", x = 1,
           y = 15,
           label = sprintf("rho = %.2f\np = %.3f", 
                          cor_O2$estimate, cor_O2$p.value),
           hjust = 0, vjust = 1, size = 3.5, fontface = "bold", color =treatment_colors["O2"]) +
  scale_color_manual(values = treatment_colors,
                     labels = c(expression(O[2]), 
                               expression(CH[4]~+~O[2])),
                     name = NULL) +
  scale_fill_manual(values = treatment_colors,
                    labels = c(expression(O[2]), 
                              expression(CH[4]~+~O[2])), 
                    name = NULL) +
  scale_linetype_manual(
    values = c("All data" = "dashed"),
    labels = c("All data"),
    name = NULL) +
  scale_x_continuous(limits = c(0, max(d15N_O2_data_2025$O2_., na.rm = TRUE) * 1.05),
                    breaks = seq(0, 15, by = 5)) +
  labs(x = expression(O[2]~("%")),
       y = expression(delta^{15}*N~("‰"))) +
  theme_bw(base_size = 10) +
  theme(
    text = element_text(family = "sans", color = "black"),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    panel.border = element_rect(color = "black", fill = NA, linewidth = 0.5),
    axis.ticks = element_line(color = "black", linewidth = 0.4),
    axis.text = element_text(color = "black", size = 10),
    axis.title = element_text(color = "black", size = 11),
    legend.position = c(0.98, 0.98),
    legend.justification = c(1, 1),
    legend.title = element_blank(),
    legend.text = element_text(size = 10),
    legend.background = element_rect(fill = "white", color = NA),
    legend.key.size = unit(5, "mm"),
    legend.spacing.y = unit(0, "mm"),
    legend.box.spacing = unit(0, "mm"),
    plot.margin = margin(5, 5, 5, 5, "mm"))+
  guides(
    color = guide_legend(order = 1, override.aes = list(shape = 21, fill = treatment_colors)),
    fill = "none",
    linetype = guide_legend(order=2)
  )

ggsave("spearman_correlation_d15N_O2.png", 
       plot = p_correlation_d15N_O2, 
       width = 140, height = 120, units = "mm", dpi = 300)

ggsave("spearman_correlation_d15N_O2.pdf",
       plot = p_correlation_d15N_O2,
       width = 140, height = 120, units = "mm",
       device = cairo_pdf,
       dpi = 300)

print(p_correlation_d15N_O2)


```

